## Bar Charts

A bar chart uses rectangular bars to visualize data. Bar charts can be displayed horizontally or vertically. The height or length of the bars are proportional to the values they represent.

Use the `barplot()` function to draw a vertical bar chart:

```{r}
x <- c("A","B","C","D")
y <- c(2,4,6,8)

barplot(y,names.arg = x,main = "Sample BarCart",col = "blue")
```

#### Example Explained

-   The `x` variable represents values in the x-axis (A,B,C,D)

-   The `y` variable represents values in the y-axis (2,4,6,8)

-   Then we use the `barplot()` function to create a bar chart of the values

-   `names.arg` defines the names of each observation in the x-axis

## Density / Bar Texture

To change the bar texture, use the `density` parameter:

```{r}
x <- c("A", "B", "C", "D")
y <- c(2, 4, 6, 8)

barplot(y, names.arg = x, density = 5)
```

## Bar Width

Use the `width` parameter to change the width of the bars:

```{r}
x <- c("A", "B", "C", "D")
y <- c(2, 4, 6, 8)

barplot(y, names.arg = x, width = c(1,2,3,4))
```

# Horizontal Bars

If you want the bars to be displayed horizontally instead of vertically, use `horiz=TRUE`:

```{r}
x <- c("A", "B", "C", "D")
y <- c(2, 4, 6, 8)

barplot(y, names.arg = x, horiz = TRUE)
```
