# R Vectors

A vector is simply a list of items that are of the same type.

To combine the list of items to a vector, use the `c()` function and separate the items by a comma.

1.  create a vector variable called **fruits**, that combine strings:

    ```{r}
    fruits <-  c("Apple","Orange","Lemon","Watermelon")

    print(fruits)
    ```

2.  create a vector marks that combines numerical values:

    ```{r}
    marks <- c(34,56,78,34,56,55)

    print(marks)
    ```

3.  create a vector **ages** with numerical values in a sequence, use the `:` operator:

    ```{r}
    ages <- (32.7:36.9)

    print(ages)
    ```

4.  create a vector of logical values:

    ```{r}
    logics <- c(TRUE,FALSE,TRUE,TRUE,TRUE,FALSE)

    print(logics)
    ```

## Vector Functions and Operations

```{r}
fruits <- c("banana", "apple", "orange", "mango", "lemon")

length(fruits)
sort(fruits)
fruits[1]
fruits[c(1, 3)]
fruits[1] <- "pear"
fruits
```

## Repeat Vectors

To repeat vectors, use the `rep()` function:

```{r}
ages <- rep(c(45,67,78),each = 2)

print(ages)
```

```{r}
nums <- rep(c(10L,34L,56L), times = 3)

print(nums)
```

```{r}
repeat_indepent <- rep(c(1,2,3), times = c(5,2,1))

repeat_indepent
```

### Generating Sequenced Vectors

To create a vector with numerical values in a sequence with the `:` operator:

```{r}
numbers <- 1:23

numbers
```

To make bigger or smaller steps in a sequence, use the `seq()` function:

```{r}
agesp <- seq(from = 10, to = 450, by = 45)


agesp
```

**Note:** The `seq()` function has three parameters: `from` is where the sequence starts, `to` is where the sequence stops, and `by` is the interval of the sequence.

# R Lists

A list in R can contain many different data types inside it. A list is a collection of data which is ordered and changeable.

To create a list, use the `list()` function:

```{r}
cars <- list("Benz","Suzuki","V8",4567)

class(cars)
cars
```

### List functions and Operations

```{r}
cars[1]
cars[3] = "Toyota"
cars
length(cars)
"Volvo" %in% cars
append("Voxy",cars)
cars
```

```{r}
append(cars, "volvo", after = 2)
cars

newlist <- cars[-1]

# Print the new list
newlist

```

## Range of Indexes

**Note:** The search will start at index 2 (included) and end at index 5 (included).

Remember that the first item has index 1.

```{r}
thislist <- list("apple", "banana", "cherry", "orange", "kiwi", "melon", "mango")

(thislist)[2:5]
```

## Join Two Lists

There are several ways to join, or concatenate, two or more lists in R.

The most common way is to use the `c()` function, which combines two elements together:

```{r}
list1 <- list("a", "b", "c")
list2 <- list(1,2,3)
list3 <- c(list1,list2)

list3
```

# R Matrices

A matrix is a two dimensional data set with columns and rows.

A column is a vertical representation of data, while a row is a horizontal representation of data.

A matrix can be created with the `matrix()` function. Specify the `nrow` and `ncol` parameters to get the amount of rows and columns:

```{r}
marksMatrix <- matrix(c(23, 56, 89, 65, 45, 56, 43, 56,78,98,65,43), nrow=4, ncol=3)
print(marksMatrix)
```

**Note:** Remember the `c()` function is used to concatenate items together.

### Matrices Functions and operations

```{r}
marksMatrix[1, 2]
marksMatrix[3,]
marksMatrix[,2]
marksMatrix[c(2,4),]
marksMatrix[,c(1,3)]

```

## Add Rows and Columns

Use the `cbind()` function to add additional columns in a Matrix:

```{r}
newMatrix <- cbind(marksMatrix,c(56,98,78,43))

newMatrix
```

**Note:** The cells in the new column must be of the same length as the existing matrix.

Use the `rbind()` function to add additional rows in a Matrix:

```{r}
nMatrix <- rbind(newMatrix,c(45,67,87,32))
nMatrix
```

## Remove Rows and Columns

Use the `c()` function to remove rows and columns in a Matrix:

```{r}
#Remove the first row and the first column
tmatrix <- nMatrix[-c(1), -c(1)]

tmatrix
```

## Number of Rows and Columns

Use the `dim()` function to find the number of rows and columns in a Matrix:

```{r}
dim(tmatrix)
```

## Matrix Length

Use the `length()` function to find the dimension of a Matrix:

```{r}
print(paste("The length is:",length(tmatrix)))

# Looping through
for (rows in 1:nrow(tmatrix)) {
  for (columns in 1:ncol(tmatrix)) {
    print(tmatrix[rows, columns])
  }
}
```

## Combine two Matrices

Again, you can use the `rbind()` or `cbind()` function to combine two or more matrices together:

```{r}
# Combine matrices
Matrix1 <- matrix(c("apple", "banana", "cherry", "grape"), nrow = 2, ncol = 2)
Matrix2 <- matrix(c("orange", "mango", "pineapple", "watermelon"), nrow = 2, ncol = 2)

# Adding it as a rows
Matrix_Combined <- rbind(Matrix1, Matrix2)
Matrix_Combined

# Adding it as a columns
Matrix_Combined <- cbind(Matrix1, Matrix2)
Matrix_Combined

```

# R Arrays

Compared to matrices, arrays can have more than two dimensions.

We can use the `array()` function to create an array, and the `dim` parameter to specify the dimensions:

```{r}
x <- c(3:18)

xarrays <- array(x,dim=c(2,4,2))
xarrays

```

#### Example Explained

In the example above we create an array with the values 1 to 24.

How does `dim=c(4,3,2)` work?\
The first and second number in the bracket specifies the amount of rows and columns.\
The last number in the bracket specifies how many dimensions we want.

**Note:** Arrays can only have one data type.

## Access Array Items

```{r}
xarrays[1,3,2]
```

The syntax is as follow: array[row position, column position, matrix level]

You can also access the whole row or column from a matrix in an array, by using the `c()` function:

```{r}
xarrays[2,1,]

cat("\n\n")

xarrays[,1,1]

cat("\n\n")
# Selecting multiple indices: xarrays[,c(1,2),1] (selects columns 1 and 2).
xarrays[,c(1,2),1]

cat("\n\n")
# Combining indices dynamically: xarrays[,c(1:3),1] (selects columns 1 through 3).
xarrays[,c(1:3),1]
```

A comma (,) before c() means that we want to access the column.

A comma (,) after c() means that we want to access the row.

# R Data Frames

Data Frames are data displayed in a format as a table.

Data Frames can have different types of data inside it. While the first column can be `character`, the second and third can be `numeric` or `logical`. However, each column should have the same type of data.

Use the `data.frame()` function to create a data frame:

```{r}
employee_data <- data.frame(
  Name = c("Alice", "Bob", "Charlie","Diana"),
  Age = c(25, 30, 35,40),
  Department = c("HR", "Finance", "IT","Marketing"),
  Is_manager = c(TRUE, FALSE, TRUE, FALSE)
)

employee_data
```

Use the `summary()` function to summarize the data from a Data Frame:

```{r}
summary(employee_data)
```

## Access Items

We can use single brackets `[ ]`, double brackets `[[ ]]` or `$` to access columns from a data frame:

```{r}
employee_data[1]

employee_data[["Department"]]

employee_data$Age
```

# R Factors

Factors are used to categorize data. Examples of factors are:

-   Demography: Male/Female

-   Music: Rock, Pop, Classic, Jazz

-   Training: Strength, Stamina

To create a factor, use the `factor()` function and add a vector as argument:

```{r}
# Create a factor
music_genre <- factor(c("Jazz", "Rock", "Classic", "Classic", "Pop", "Jazz", "Rock", "Jazz"))

# Print the factor
music_genre
```

You can see from the example above that that the factor has four levels (categories): Classic, Jazz, Pop and Rock.

To only print the levels, use the `levels()` function:

```{r}
music_genre <- factor(c("Jazz", "Rock", "Classic", "Classic", "Pop", "Jazz", "Rock", "Jazz"))

levels(music_genre)
```

You can also set the levels, by adding the `levels` argument inside the `factor()` function:

```{r}
music_genre <- factor(c("Jazz", "Rock", "Classic", "Classic", "Pop", "Jazz", "Rock", "Jazz"), levels = c("Classic", "Jazz", "Pop", "Rock", "Other"))

levels(music_genre)
```
