## Line Graphs

A line graph has a line that connects all the points in a diagram.

To create a line, use the `plot()` function and add the `type` parameter with a value of `"l"`

## Line Width

To change the width of the line, use the `lwd` parameter (`1` is default, while `0.5` means 50% smaller, and `2` means 100% larger):

```{r}
plot(1:13,xlab="X axis",ylab="Y axis",col="red",lwd=8,type="l")
```

## Line Styles

The line is solid by default. Use the `lty` parameter with a value from **0 to 6** to specify the line format.

Available parameter values for `lty`:

-   `0` removes the line

-   `1` displays a solid line

-   `2` displays a dashed line

-   `3` displays a dotted line

-   `4` displays a "dot dashed" line

-   `5` displays a "long dashed" line

-   `6` displays a "two dashed" line

For example, `lty=3` will display a dotted line instead of a solid line:

```{r}
plot(1:20,11:30,type="l",col="orange",lwd=5,lty=3)
```

## Multiple Lines

To display more than one line in a graph, use the `plot()` function together with the `lines()` function:

```{r}
line1 <- 0:30
line2 <- 10:40

plot(line1,type="l",col="red",lwd=2,lty=3)
lines(line2,type="l",col="blue",lwd=4,lty=2)
```
