# Pie Charts

A pie chart is a circular graphical view of data.

Use the `pie()` function to draw pie charts:

```{r}
shares <- c(10,40,20,30)

pie(shares)
```

By default, the plotting of the first pie starts from the x-axis and move **counterclockwise**.

**Note:** The size of each pie is determined by comparing the value with all the other values, by using this formula:

The value divided by the sum of all values: `x/sum(x)`

## Start Angle

You can change the start angle of the pie chart with the `init.angle` parameter.

The value of `init.angle` is defined with angle in degrees, where default angle is 0.

```{r}
ages <- c(23,54,11,7,5)

pie(ages,init.angle=90)
```

## Labels and Header

Use the `label` parameter to add a label to the pie chart, and use the `main` parameter to add a header:

```{r}
fruits <- c(10,50,20,5,15)
flabels <- c("Apples","Grapes","Oranges","Melons","Lemons")

pie(fruits,main="Fruits Distribution",label=flabels,init.angle=90)
```

## Colors

You can add a color to each pie with the `col` parameter:

```{r}
share <- c(50,20,15,15)
slabels <- c("Brian","Lucky","Trinity","Tevin","Tyrus")
scolors <- c("blue","red","orange","green","grey")

pie(share,main="Company Shares",label = slabels,col=scolors,init.angle = 90)
```

## Legend

To add a list of explanation for each pie, use the `legend()` function:

```{r}
share <- c(50,20,15,15)
slabels <- c("Brian","Lucky","Trinity","Tevin","Tyrus")
scolors <- c("blue","red","orange","green","grey")

pie(share,main="Company Shares",col=scolors,init.angle = 90)


# Adding legend
legend("topleft",slabels,fill=scolors,title="key")
```

The legend can be positioned as either:

`bottomright`, `bottom`, `bottomleft`, `left`, `topleft`, `top`, `topright`, `right`, `center`
