### **Comprehensive R Programming Task**

#### **Scenario: Analyzing Student Grades**

You are tasked with creating an R script to process and analyze student grade data for a class of 10 students. Follow these steps:

#### **Step 1: Variables and Vectors**

1.  Create a vector `students` with the names of 10 students (e.g., "Alice", "Bob", etc.).

    ```{r}
    students <- c("Alice","Bob","Marley","Rick","Ross","Navy","Kenzo","Rich","Mavoko","Jose")
    students
    ```

2.  Create a vector `math_grades` with random numbers between 50 and 100, representing each student’s math grade.

    ```{r}
    math_grades <- as.integer(c(runif(10,min = 50,max = 100)))
    math_grades
    ```

3.  Create another vector `science_grades` with random numbers between 50 and 100, representing each student’s science grade.

    ```{r}
    science_grades <- as.integer(c(runif(10,min = 50,max = 100)))
    science_grades
    ```

#### **Step 2: Lists**

4.  Combine `students`, `math_grades`, and `science_grades` into a list called `student_data`.

    ```{r}
    student_data <- list(students,math_grades,science_grades)
    student_data
    ```

#### **Step 3: Matrices**

5.  Create a matrix called `grades_matrix` with 10 rows and 2 columns using `math_grades` and `science_grades`. Label the rows with student names and the columns as "Math" and "Science."

    ```{r}
    grades_matrix <- matrix(c(math_grades,science_grades),nrow = 10,ncol = 2)
    colnames(grades_matrix) <- c("Math","Science")
    grades_matrix
    ```

#### **Step 4: Data Frames**

6.  Convert `grades_matrix` into a data frame called `grades_df` with columns "Student", "Math", and "Science."

    ```{r}
    grades_df <- data.frame(
      Student <- students,
      Math <- math_grades,
      Science <- science_grades
    )

    colnames(grades_df) <- c("Student","Math","Science")

    grades_df
    ```

7.  Add a new column to `grades_df` called `Average`, which contains the average of each student’s math and science grades.

    ```{r}
    Average <- (grades_df$Math + grades_df$Science)/2
    grades_df <- cbind(grades_df,Average)
    grades_df

    ```

#### **Step 5: Conditional Statements and Loops**

8.  Write a `for` loop that prints each student’s name and whether they passed or failed each subject. A passing grade is 60 or higher.

    ```{r}
    for(i in 1:nrow(grades_df)){
      name <- grades_df$Student[i]
      math <- grades_df$Math[i]
      science <- grades_df$Science[i]
      
      if(math >= 60){
        mstate = "Pass"
      }else{
        mstate = "Fail"
      }
      
      if(science >= 60){
        sstate = "Pass"
      }else{
        sstate = "Fail"
      }
      
      cat("Name:",name,"Math:",mstate,"Science:",sstate,"\n")
    }
    ```

9.  Write a function `calculate_grade_category` that takes a numeric grade as input and returns:

    -   "Fail" if the grade is below 60,

    -   "Pass" if the grade is between 60 and 79,

    -   "Distinction" if the grade is 80 or above.

        ```{r}
        calculate_grade_category <- function(grade){
          if(grade < 60){
            "Fail"
          }else if(grade >= 60 & grade <= 79){
            "Pass"
          }else if(grade >= 80){
            "Distinction"
          }
        }
        ```

10. Use `sapply()` to apply `calculate_grade_category` to the "Average" column in `grades_df` and create a new column called `Grade_Category`.

    ```{r}
    grades_df$Grade_Category <- sapply(grades_df$Average, calculate_grade_category)
    ```

#### **Step 6: String Manipulation**

11. Update the "Student" column in `grades_df` to include the phrase "(Top Student)" next to any student with an average above 90.

    ```{r}
    grades_df$Student <- ifelse(grades_df$Average > 90, paste(grades_df$Student, "(Top Student)"), grades_df$Student)
    ```

#### **Step 7: File Input/Output**

12. Save the `grades_df` to a CSV file named `student_grades.csv`.

    ```{r}
    write.csv(grades_df, "student_grades.csv", row.names = FALSE)  # Save to CSV
    ```

13. Read the file back into R as `grades_df_imported` and print the first few rows to confirm the data was saved and loaded correctly.

    ```{r}
    grades_df_imported <- read.csv("student_grades.csv")  # Read from CSV
    print(head(grades_df_imported))  # Print first few rows
    ```

#### **Step 8: Plotting**

14. Create a bar plot of the average grades for all students. Customize it with:

    -   A title ("Average Grades of Students"),

    -   Labeled axes ("Student Names" and "Average Grade"),

    -   A different color for each bar.

        ```{r}
        barplot(grades_df$Average, names.arg = grades_df$Student, col = rainbow(10), main = "Average Grades of Students", xlab = "Student Names", ylab = "Average Grade", las = 2)
        ```

15. Create a scatter plot of math grades vs. science grades. Add:

    -   A title ("Math vs. Science Grades"),

    -   A linear trend line,

    -   Labels for each student.

        ```{r}
        plot(grades_df$Math, grades_df$Science, main = "Math vs. Science Grades", xlab = "Math Grade", ylab = "Science Grade", pch = 19, col = "blue")
        abline(lm(Science ~ Math, data = grades_df), col = "red")  # Add trend line
        ```

#### **Step 9: Bonus Challenge**

16. Write a function `get_top_students` that:

    -   Takes `grades_df` as input,

    -   Returns a data frame containing only students with an average grade of 80 or above,

    -   Plots a pie chart of their grade categories.

        ```{r}
        get_top_students <- function(grades_df) {
          top_students <- subset(grades_df, Average >= 80)
          pie(table(top_students$Grade_Category), main = "Grade Category Distribution of Top Students")
          return(top_students)
        }

        top_students <- get_top_students(grades_df)  # Get top students
        print(top_students)  # Print top students
        ```
